```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Overview

`metar` combines several utilities, including
many packages of the core [tidyverse](https://www.tidyverse.org/),
the [sticky](https://CRAN.R-Project.org/package=sticky) package for persistent attributes,
and the [data.table](https://CRAN.R-Project.org/package=data.table) package for fast I/O,
to provide a comprehensive interface for working with data and its metadata.
Currently the package provides utilities for:

- Reading from CSVY files, which provide a convenient way of storing tabular data and its metadata together while still being readable by common CSV-reading utilities.
- Adding metadata to in-memory `data.frame`s with a `dplyr`-like syntax

# Installation

This package is not on CRAN (yet!), but is easy enough to install from GitHub:

```{r, eval = FALSE}
devtools::install_github("ashiklom/metar")
```

# Usage

Add metadata directly to objects:

```{r}
library(metar)

obj <- rnorm(100)
metadata(obj) <- list(title = "Normal draws", description = "Some random normal draws")
metadata(obj)
metadata(obj, "title", "description")
```

You can also use the pipe-friendly `add_metadata`:

```{r}
obj2 <- runif(100) %>%
  add_metadata(title = "Normal draws") %>%
  # Metadata can be added or overwritten
  add_metadata(author = "RNG", title = "Just kidding, they're uniform draws") %>%
  # Metadata can be any R object
  add_metadata(range = list(min = 0, max = 1))
metadata(obj2)
metadata(obj2, "range")
```

Add metadata to each column of the `iris` dataset:

```{r}
library(dplyr)

iris_md <- iris %>%
  as_tibble() %>%
  add_column_metadata(
    # Metadata for individual columns
    Sepal.Length = list(title = "Sepal length", unit = "cm"),
    Sepal.Width = list(title = "Sepal width", unit = "in"),
    # .root sets metadata for the data frame as a whole
    .root = list(description = "The famous Iris dataset", author = "Not Alexey Shiklomanov")
  )
```

Metadata are stored as `attributes`.

```{r}
attr(iris_md, "description")
attr(iris_md, "author")
attr(iris_md$Sepal.Length, "title")
attributes(iris_md$Sepal.Width)
```

These attributes are made persistent through many operations by the `sticky` package.

```{r}
metadata(iris_md)
metadata(iris_md[1, ], "description")
metadata(iris_md$Sepal.Length[1:5], "title")
```

The CSVY format is just a CSV file with a YAML header.

```{r}
system.file("examples/example1.csvy", package = "metar") %>%
  readLines() %>%
  writeLines()
```

If read in with the `read_csvy` function, all metadata will be automatically applied:

```{r}
dat <- read_csvy(system.file("examples/example1.csvy", package = "metar"))
dat
metadata(dat)
lapply(dat, metadata)
```

Note that because `var2` was specified as an `integer` class, it was also read in as an `integer` rather than a `numeric`.

```{r}
class(dat$var2)
```
